Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-08 files/Chapter3/femconvergence.tex
Veranderinge 10-06
symbols

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Convergence of eigenvalues in Finite Element Analysis}
In this section, the work in the textbook \cite[p.228-236]{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for clarity.\\

@@

\subsection{General eigenvalue problem}
Find a function $u \in H$ and number $\lambda \in \mathcal{R}$ such that
\begin{equation}
	b(u,v) = \lambda (u,v) \label{GEVP}
\end{equation} for some $v \in V$.

@@

\textbf{Rayleigh quotient}
\begin{equation*}
	R(v) = \frac{b(v,v)}{(v,v)} \ \text { for } v \in V.
\end{equation*}

@@

Consider a interpolation operator $\Pi$. This projection is linear, i.e.
\begin{align*}
	\Pi(f + g) & = \Pi f + \Pi g,\\
	\Pi(\alpha f) & = \alpha \Pi f \ \text{ for a constant } \alpha. 
\end{align*}

@@

The following notation is introduced.
\begin{itemize}
	\item[] $\mathcal{P}_j(I_e)$: Is the set of all polynomials on the interval $I_e$ of degree at most $j$.
	\item[] $r(\Pi_e)$: If the range of $\Pi_e$ is contained in $\mathcal{P}_j(I_e)$ and $k<j$ is the largest integer such that $\Pi_e f = f$ for each $f \in \mathcal{P}_j(I_e)$, then $r(\Pi_e) = k$.
	\item[] $s(\Pi_e)$: Is a integer and the largest order derivative used in the definition of $\Pi_e$.
\end{itemize}

@@

For the requirements of this dissertation, this can be simplified with assumptions. The first assumption is a reiteration that the eigenvalue problem is one-dimensional. Another assumption is that the basis of $S^h$ consists of polynomials. With these assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^(m) - u^(m)|| \leq C h^{k-m} || u^(k)|| \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

@@

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_i \in V | i = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_i \in V | i = 1,2,...,m\right\}$.\\

@@

The space $S^h$ is a finite dimensional subspace of $V$. Therefore any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}

@@

\newtheorem{Prop_1}{Proposition}
\begin{Prop_1}
 $\mu_{m} > 0$ if and only if $\dim S_{m} = m$.
\end{Prop_1}
\begin{proof}
	To show that the dimension of $S_m = m$, suppose that the elements of $B_m$ are linearly dependent. Then there exists a $u \in B_m$ such that $Pu = 0$ and consequently $\mu_m = 0$. The result follows from the contra-positive.
\end{proof}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Convergence of eigenvalues and eigenfunctions}
In this section, the work in the textbook \cite[p.228-236]{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for clarity.\\

@@

\subsection{General eigenvalue problem}
Find a function $u \in H$ and number $\lambda \in R$ such that
\begin{equation}
	b(u,v) = \lambda (u,v) \label{GEVP}
\end{equation} for some $v \in V$.

@@

\textbf{Rayleigh quotient}
\begin{equation*}
	R(v) = \frac{b(v,v)}{(v,v)} \ \text { for } v \in V.
\end{equation*} \label{sym:Rayleigh}

@@

Consider a interpolation operator $\Pi$.\label{sym:interpolation} This projection is linear, i.e.
\begin{align*}
	\Pi(f + g) & = \Pi f + \Pi g,\\
	\Pi(\alpha f) & = \alpha \Pi f \ \text{ for a constant } \alpha. 
\end{align*}

@@

The following notation is introduced.
\begin{itemize}
	\item[] $\mathcal{P}_j(I_e)$: Is the set of all polynomials on the interval $I_e$ of degree at most $j$. \label{sym:poly}
	\item[] $r(\Pi_e)$: If the range of $\Pi_e$ is contained in $\mathcal{P}_j(I_e)$ and $k<j$ is the largest integer such that $\Pi_e f = f$ for each $f \in \mathcal{P}_j(I_e)$, then $r(\Pi_e) = k$.
	\item[] $s(\Pi_e)$: Is a integer and the largest order derivative used in the definition of $\Pi_e$.
\end{itemize}

@@

For the requirements of this dissertation, this can be simplified with assumptions. The first assumption is a reiteration that the eigenvalue problem is one-dimensional. Another assumption is that the basis of $S^h$ consists of polynomials. With these assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}|| \leq C h^{k-m} || u^{(k)}|| \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

@@

Fix $m \in \mathbb{N}$. \label{sym:natural} Let $\left\{ \phi_i \in V | i = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_i \in V | i = 1,2,...,m\right\}$.\\

@@

The space $S^h$\label{sym:Sh1} is a finite dimensional subspace of $V$. Therefore any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}

@@

\newtheorem{Prop_1}{Proposition}
\begin{Prop_1}
 $\mu_{m} > 0$ if and only if $\dim S_{m} = m$. \label{sym:mu}
\end{Prop_1}
\begin{proof}
	To show that the dimension of $S_m = m$, suppose that the elements of $B_m$ are linearly dependent. Then there exists a $u \in B_m$ such that $Pu = 0$ and consequently $\mu_m = 0$. The result follows from the contra-positive.
\end{proof}

### 2023-11-08 table_of_symbols.tex
symbols

The following table provides a list of symbols that will be used throughout this dissertation along with their descriptions:

\noindent
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\endhead % Header to be repeated on every page
\( R^n \) & n-dimensional space of real numbers & \pageref{sym:Rn} \\
\( u/w \) & displacement vector & \pageref{sym:u} \\
\( \Omega \) & subset of $R^n$ & \pageref{sym:Omega} \\
\( \rho \) & density & \pageref{sym:rho} \\
\( \partial_x^n \) & n-th partial derivative with respect to $x$ & \pageref{sym:partial_diff} \\
\( \sigma_{ij} \) & element of the stress tensor T & \pageref{sym:sigmaij} \\
\( Q \) & force per unit volume & \pageref{sym:Q} \\
\( \textrm{div}T \) & divergence of T & \pageref{sym:divT} \\
\( \textrm{Tr}(T) \) & Trace of T & \pageref{sym:TrT} \\
\( \mathcal{E} \) & infinitesimal strain tensor & \pageref{sym:mathcalE} \\
\( \varepsilon_{ij} \) & element of the infinitesimal strain tensor $\mathcal{E}$ & \pageref{sym:varepsilonij} \\
\( E \) & Young's modulus & \pageref{sym:E} \\
\( \nu \) & Poisson's ratio & \pageref{sym:nu} \\
\( t \) & time & \pageref{sym:t} \\
\( G \) & shear modulus of elasticity & \pageref{sym:G} \\
\( \tau \) & dimensionless time & \pageref{sym:tau} \\
\( \xi \) & dimensionless space & \pageref{sym:xi} \\
\( Q^* \) & dimensionless force per unit volume & \pageref{sym:Qstar} \\
\( \sigma^*_{ij} \) & dimensionless stress tensor element & \pageref{sym:sigmaijstar} \\
\( \ell \) & dimension representing length & \pageref{sym:ell} \\
\( \kappa^2 \) & some dimensionless constant/shear correction factor & \pageref{sym:kappa2}/\pageref{sym:kappa2T} \\
\( I \) & identity matrix & \pageref{sym:I} \\
\( \gamma \) & a dimensionless constant & \pageref{sym:gamma} \\
\( n \) & a normal vector & \pageref{sym:n} \\
\( \Sigma/\Gamma \) & distinct parts of $\Omega$ & \pageref{sym:SigmaGamma} \\
\( \phi \) & arbitrary vector/rotation of cross-section of Timoshenko beam & \pageref{sym:phi}/\pageref{sym:phiT} \\
\( C^n \) & set of n-times continuously differentiable functions & \pageref{sym:Cn} \\
\( a(\cdot,\cdot),b(\cdot,\cdot),c(\cdot,\cdot) \) & bilinear forms & \pageref{sym:a},\pageref{sym:b},\pageref{sym:c} \\
\( T(\Omega) \) & test function space on $\Omega$ & \pageref{sym:TOmega} \\
\( dV \) & volume integral meassure & \pageref{sym:dV} \\
\( dS \) & surface integral meassure & \pageref{sym:dS} \\
\( dA \) & area integral meassure & \pageref{sym:dA} \\
\( ds \) & line integral meassure & \pageref{sym:ds} \\
\( A \) & area of a cross-section & \pageref{sym:A} \\
\( V \) & shear force & \pageref{sym:V} \\
\( I \) & area moment of inertia & \pageref{sym:Iinertia} \\
\( M \) & moment & \pageref{sym:M} \\
\( \alpha/\beta \) & dimensionless constants & \pageref{sym:alphabeta} \\
\( e_n \) & unit vector in the direction n & \pageref{sym:en} \\
\( \mathbf{Q} \) & constitutive equation for the Reissner-Mindlin plate model & \pageref{sym:bfQ} \\
\( L^n \) & space of n-power Lebesgue integrable functions & \pageref{sym:Ln} \\
\( H^n \) & space of functiontions with weak derivatives up to order n (n'th dimensional Sobolev space) & \pageref{sym:Hn} \\
\( X \) & global space & \pageref{sym:Xspace} \\
\( W \) & inertia space & \pageref{sym:Wspace} \\
\( V \) & energy space & \pageref{sym:Vspace} \\
\( ||\cdot||_X \) & norm in the space $X$ & \pageref{sym:norm} \\
\( C^\infty \) & space of smooth functions & \pageref{sym:Cinfty} \\
\( C^\infty_0 \) & space of smooth functions with compact support & \pageref{sym:Cinftyzero} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\

>>>>

% Vector Spaces and Related Concepts
\noindent
\textbf{Vector Spaces and Related Concepts}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( \mathbb{R}^n \) & n-dimensional space of real numbers & \pageref{sym:Rn} \\
\( \Omega \) & subset of \( \mathbb{R}^n \), representing a body/reference configuration & \pageref{sym:Omega} \\
\( C^n \) & set of n-times continuously differentiable functions & \pageref{sym:Cn} \\
\( L^n \) & space of n-power Lebesgue integrable functions & \pageref{sym:Ln} \\
\( H^n \) & space of functions with weak derivatives up to order n (n'th dimensional Sobolev space) & \pageref{sym:Hn} \\
\( C^\infty \) & space of smooth functions & \pageref{sym:Cinfty} \\
\( C^\infty_0 \) & space of smooth functions with compact support & \pageref{sym:Cinftyzero} \\
\( \mathcal{P}_j \) & set of all polynomials of degree at most \( j \) & \pageref{sym:poly} \\
\( \mathbb{N} \) & set of natural numbers \( j \) & \pageref{sym:natural} \\
\( S^h \) & finite dimensional subspace & \pageref{sym:Sh1} \\

% Mathematical Measures and Operations
\noindent
\textbf{Mathematical Measures and Operations}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( \partial_x^n \) & n-th partial derivative with respect to \( x \) & \pageref{sym:partial_diff} \\
\( \textrm{div} T \) & divergence of \( T \) & \pageref{sym:divT} \\
\( \textrm{Tr}(T) \) & Trace of \( T \) & \pageref{sym:TrT} \\
\( dV \) & volume integral measure & \pageref{sym:dV} \\
\( dS \) & surface integral measure & \pageref{sym:dS} \\
\( dA \) & area integral measure & \pageref{sym:dA} \\
\( ds \) & line integral measure & \pageref{sym:ds} \\
\( a(\cdot,\cdot), b(\cdot,\cdot), c(\cdot,\cdot) \) & bilinear forms & \pageref{sym:a}, \pageref{sym:b}, \pageref{sym:c} \\
\( ||\cdot||_X \) & norm in the space \( X \) & \pageref{sym:norm} \\
\( T(\Omega) \) & test function space on \( \Omega \) & \pageref{sym:TOmega} \\
\( \mathcal{E}(\cdot) \) & energy function & \pageref{sym:Energy} \\
\( Pu \) & Projection of the vector u & \pageref{sym:Pu} \\
\( R(\cdot) \) & Rayleigh quotient & \pageref{sym:Rayleigh} \\
\( \Pi \) & interpolation operator & \pageref{sym:interpolation} \\

% Physical Quantities and Parameters
\noindent
\textbf{Physical Quantities and Parameters}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( u \) or \( w \) & displacement vector & \pageref{sym:u} \\
\( \rho \) & density & \pageref{sym:rho} \\
\( \sigma_{ij} \) & element of the stress tensor \( T \) & \pageref{sym:sigmaij} \\
\( Q \) & force per unit volume & \pageref{sym:Q} \\
\( \mathcal{E} \) & infinitesimal strain tensor & \pageref{sym:mathcalE} \\
\( \varepsilon_{ij} \) & element of the infinitesimal strain tensor \( \mathcal{E} \) & \pageref{sym:varepsilonij} \\
\( E \) & Young's modulus & \pageref{sym:E} \\
\( \nu \) & Poisson's ratio & \pageref{sym:nu} \\
\( t \) & time & \pageref{sym:t} \\
\( G \) & shear modulus of elasticity & \pageref{sym:G} \\
\( \tau \) & dimensionless time & \pageref{sym:tau} \\
\( \xi \) & dimensionless space & \pageref{sym:xi} \\
\( Q^* \) & dimensionless force per unit volume & \pageref{sym:Qstar} \\
\( \sigma^*_{ij} \) & dimensionless stress tensor element & \pageref{sym:sigmaijstar} \\
\( \ell \) & dimension representing length & \pageref{sym:ell} \\
\( b \) &  dimension representing width & \pageref{sym:width} \\
\( h \) & dimension representing height & \pageref{sym:height} \\
\( \kappa^2 \) & some dimensionless constant/shear correction factor & \pageref{sym:kappa2} \\
\( I \) & identity matrix & \pageref{sym:I} \\
\( \gamma \) & a dimensionless constant & \pageref{sym:gamma} \\
\( n \) & a normal vector & \pageref{sym:n} \\
\( \Sigma/\Gamma \) & distinct parts of \( \Omega \) & \pageref{sym:SigmaGamma} \\
\( \phi \) & arbitrary vector/rotation of cross-section of Timoshenko beam & \pageref{sym:phi} \\
\( A \) & area of a cross-section & \pageref{sym:A} \\
\( V \) & shear force & \pageref{sym:V} \\
\( I \) & area moment of inertia & \pageref{sym:Iinertia} \\
\( M \) & moment & \pageref{sym:M} \\
\( \alpha/\beta \) & dimensionless constants & \pageref{sym:alphabeta} \\
\( \lambda \) & eigenvalue & \pageref{sym:lambda} \\
\( \mu \) & eigenvector & \pageref{sym:mu} \\
